---
title       : Hawaii Electricity Prices
subtitle    : 2008-2012
author      : Michael Ching
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Electricity Prices in Hawaii
1. Prices for electricity in Hawaii are among the highest in the United States
1. These prices vary not only by time but by island and sector
1. Sectors include Residential, Commercial, and Street Light uses
1. The following figure shows prices from 2008 to 2012 plotted with `ggplot2` (Source: [Hawaii Open Data](https://data.hawaii.gov/Government-Wide-Support/HI-Electricity-Prices/74g9-vewt))

--- .class #id 

## Price Change by Island and Sector over Time

```{r, echo=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
dat <- read.csv("~/Dropbox/Mike/Coursera/9DevelopingDataProducts/HI_electricity_prices/data/prices.csv")
dat <- dat[dat$Sector !=  "All Sectors", ]
dat$RptMonth <- as.Date(strptime(dat$RptMonth, format = "%m/%d/%Y"))
ggplot(dat, aes(x = RptMonth, y = Price, group = Island, color = Island)) + 
    geom_line() + facet_grid(Sector ~ .) + ylab("Price (cents per kwh)") +
    xlab("Time")
```

---

## Exploring Price Variation
I created a [Shiny app](https://mching.shinyapps.io/HI_electricity_prices/) to provide medians for the price by time, island and sector

1. This is an example of how the app works. 
1. We can give it dates such as 2010-01-01 through 2010-12-01 and ask for the island of Oahu and sector of Residential rates. 
1. The app returns the following value in cents per kilowatt-hour. 
```{r, echo=FALSE}
Island_input <- "Oahu"
Sector_input <- "Residential"
start_date <- as.Date("2010-01-01")
end_date <- as.Date("2010-12-01")
filter(dat, Island == Island_input & Sector == Sector_input) %>%
    filter(RptMonth >= start_date & RptMonth <= end_date) %>%
    select(Price) %>%
    summarise_each(funs(median))
```


---

## Conclusion

This application makes it easy to extract median electricity prices for a given island, sector and dates.

