---
title: "Hawai'i Electricity Rates Project Documentation"
author: "Michael Ching"
date: "February 18, 2015"
output: html_document
keep_md: true
---

## Usage
### Summary
This simple application allows you to find information on median electricity prices in the state of Hawai'i. You select a range of dates, the sector using the electricity, and the island. The application returns the median value for the given inputs.

### Date Range
Dates may be selected between 30 January 2008 and 31 March 2012

### Sector
Possible sectors include 

1. Commercial (default)
1. Residential
1. Street Light

### Island
Available islands are

1. Hawaii (default)
1. Kauai
1. Lanai
1. Maui
1. Molokai
1. Oahu

### Output
The output result is provided in cents per kilowatt-hour.

## Methods
Data were obtained from the Hawai'i Open Data initiative server at https://data.hawaii.gov/Government-Wide-Support/HI-Electricity-Prices/74g9-vewt.

### Download
```{r, cache=TRUE}
library(RCurl)
data_url <- "https://data.hawaii.gov/api/views/74g9-vewt/rows.csv?accessType=DOWNLOAD"
timestamp()
x <- getURL(data_url)
dat <- read.csv(textConnection(x))
write.csv(dat, file = "./data/prices.csv")
```

### Processing
```{r}
dat <- read.csv(file = "./data/prices.csv")
```

Data were then processed to convert the month variable into R's Date class. 
```{r}
dat$RptMonth <- as.Date(strptime(dat$RptMonth, format = "%m/%d/%Y"))
```

The dataset contains information on electricity rates per sector. The three different sectors are commercial, residential and street lights. There is also another constructed summary rate for All Sectors, and I took this out of the dataset.

```{r}
levels(dat$Sector)
dat <- dat[dat$Sector !=  "All Sectors", ]
```

### Median Calculation Script
The R package `dplyr` was used to manipulate the data to find the median for given inputs. Below is an example of how the script works with some sample inputs.
```{r}
library(dplyr)

Island_input <- "Oahu"
Sector_input <- "Residential"
start_date <- as.Date("2010-01-01")
end_date <- as.Date("2010-12-01")

filter(dat, Island == Island_input & Sector == Sector_input) %>%
    filter(RptMonth >= start_date & RptMonth <= end_date) %>%
    select(Price) %>%
    summarise_each(funs(median))
```

